---
title: "Hands-On Programming with R"
author: "Guilber"
date: 2026-07-14
categories: [Book, R, Programming]
---

## References

Grolemund, G. (2014). *Hands-On Programming with R: Write Your Own Functions and Simulations*. O'Reilly Media.

- Book site: <https://rstudio-education.github.io/hopr/>

tidyquant: Tidy Quantitative Financial Analysis by [Matt Dancho](https://github.com/mdancho84)

- Site: <https://business-science.github.io/tidyquant/index.html>

## Purpose

Practice the concepts in this book using datasets obtained with the tidyquant library.

## Quick Overview of TidyQuant

`tidyquant` is an R package that integrates quantitative financial analysis with the tidyverse, it pulls financial/economic data via `quantmod`, `TTR` and `xst` and let us work with it using tidy, pipeable workflows for charting, technical analysis and portfolio/performance analytics.

### 1. Install

```{r}
#| output: false
install.packages("tidyquant")
```

### 2. Load it

```{r}
#| output: false
library(tidyquant)
```

This also loads core tidyverse pieces (`dplyr`, `ggplot2`, `tidyr`, `purrr`, `lubridate`) plus `xts`, `quantmod`, `TTR`, `PerformanceAnalytics`.

### 3. Pull data with `tq_get()`














## Notes

### Chapter 1 — The Very Basics, practiced with tidyquant

**1. Objects & assignment (`<-`)**

Store a ticker and a data pull as named objects, just like `die <- 1:6` in
the book:

```{r}
ticker <- "AAPL"
prices <- tq_get(ticker, get = "stock.prices", from = "2023-01-01")
```

`ticker` and `prices` now live in the environment and can be reused below.

**2. Vectors & element-wise operations**

`prices$close` is just a numeric vector. R applies arithmetic element-wise,
the same way `die + 1` adds 1 to every face:

```{r}
daily_change <- prices$close - lag(prices$close)
pct_change <- round(daily_change / lag(prices$close) * 100, 2)
```

**3. Calling functions with arguments**

`tq_get()` above is a function call with named arguments (`get`, `from`),
same idea as `sample(1:6, size = 1)` or `round(x, digits = 2)` in the book.
Defaults matter too — leaving out `from` would just pull the default
history length.

**4. Writing a user-defined function**

Wrap the percent-change logic from step 2 into a small function, mirroring
the book's `roll()` example (name, body, arguments, return value = last
line):

```{r}
pct_change_close <- function(data, digits = 2) {
  change <- data$close - lag(data$close)
  round(change / lag(data$close) * 100, digits)
}

pct_change_close(prices)
```

### Takeaways

- Objects in tidyquant workflows are usually data frames, but the
  underlying columns are still plain vectors — Chapter 1's rules apply
  directly.
- `tq_get()` is the recurring "function call with arguments" pattern that
  will show up throughout the rest of the book's exercises.
