---
title: "Hands-On Programming with R"
author: "Guilber"
date: 2026-07-14
categories: [Book, R, Programming]
---

## References

Grolemund, G. (2014). *Hands-On Programming with R: Write Your Own Functions and Simulations*. O'Reilly Media.

- Book site: <https://rstudio-education.github.io/hopr/>

tidyquant: Tidy Quantitative Financial Analysis by [Matt Dancho](https://github.com/mdancho84)

- Site: <https://business-science.github.io/tidyquant/index.html>

## Purpose

Practice the concepts in *Hands-On Programming with R* using datasets obtained with the `tidyquant` library.

## Quick Overview of TidyQuant

`tidyquant` is an R package that integrates quantitative financial analysis with the tidyverse, it pulls financial/economic data via `quantmod`, `TTR` and `xst` and let us work with it using tidy, pipeable workflows for charting, technical analysis and portfolio/performance analytics.

### 1. Install

```{r}
#| output: false
install.packages("tidyquant")
```

### 2. Load it

```{r}
#| output: false
library(tidyquant)
```

This also loads core tidyverse pieces (`dplyr`, `ggplot2`, `tidyr`, `purrr`, `lubridate`) plus `xts`, `quantmod`, `TTR`, `PerformanceAnalytics`.

### 3. Pull some data

```{r}

# See what indexes are available
tq_index_options()

# Pull the full S&P 500 constituent list
sp500 <- tq_index("SP500")
head(sp500)

```

This is all what we will need for this practice.

## Practice

### Chapter 2 - The Very Basics

I'm going to skip the very very basics, like choose an editor and set it up...

**1. Objects & assignment (`<-`)**

To create an R object, choose a name and then use the less-than symbol, <, followed by a minus sign, -, to save data into it. Using the previous code,

```{r}
#| output: false
sp500 <- tq_index("SP500")
# sp500 is the object that we create to store the pulled data
names(sp500)
```

**2. Vectors & element-wise operations**

Well, what is a vector? In mathematics, a vector is an object that has both magnitude (size/length) and direction. It's typically represented as an arrow, or numerically as an ordered list (tuple) of numbers called components or coordinates.

In R this has more to do with an ordered list, using the same tibble that we use before `sp500$symbol` is just a numeric vector for R.

```{r}
# We can show all the elements just running
# sp500$weight
# But to save space we will use head() function
head(sp500$weight, 10) 

```
this shows us the first 10 elements in the vector `sp500$weight`.

Now we will run an aritmetic operatin with this vector. When R performs element-wise execution, it matches up vectors and then manipulates each pair of elements independently. We will do `sp500$weight + 1` and this works 

  0.07886066 + 1 = 1.07886066

  0.07370111 + 1 = 1.07370111
  
  0.04503111 + 1 = 1.04503111
  
  0.03824187 + 1 = 1.03824187
  
  ...

```{r}
# Here R will performs element-wise execution
# This adds 1 to each element in the vector sp500$weight
head(sp500$weight + 1, 10) 
```

**2. Functions**
